---
title: "q1_analysis"
author: "Chhiring Lama"
date: "2026-04-27"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
header-includes:
  - \usepackage{fancyhdr}
  - \usepackage{lipsum}
  - \pagestyle{fancy}
  - \fancyhead[R]{\thepage}
  - \fancypagestyle{plain}{\pagestyle{fancy}}
  - \usepackage{enumitem}
  - \usepackage{float}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gtsummary)
library(gt)
library(corrplot)
library(caret)
library(naniar)
library(mgcv)
library(kableExtra)
library(sjPlot)
library(performance)
library(DHARMa)
library(patchwork)
library(tidyr)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

options(scipen = 999)
```

### Import and clean dataset
```{r}
pancreatic_df <- read_csv("data/ebl.csv") |> 
  janitor::clean_names() |> 
  mutate(sex = case_when(sex == 0 ~ "Male", 
                         sex == 1 ~ "Female"), 
         sex = fct_relevel(sex, "Male"),
         race2 = case_when(is.na(race) == TRUE ~ "NA", 
                         TRUE ~ "non-NA"), 
         race = case_when(race == 0 ~ "White",
                          race == 1 ~ "Black", 
                          race == 2 ~ "Asian",
                          race == 3 ~ "Hispanic"),
         race = fct_relevel(race, "White"), 
         op = as.factor(case_when(op == 0 ~ "Body or tail", 
                        op == 1 ~ "Whipple operation")),
         ven_res = as.factor(case_when(ven_res == 0 ~ "No", 
                             ven_res == 1 ~ "Yes")), 
         neo_chemo = as.factor(case_when(neo_chemo == 0 ~ "No", 
                             neo_chemo == 1 ~ "Yes")), 
         neo_xrt = as.factor(case_when(neo_xrt == 0 ~ "No", 
                             neo_xrt == 1 ~ "Yes")))

#head(pancreatic_df)
```

table 1: demographic characteristics + others
a) Grouped by Sex
```{r}
table1  <- pancreatic_df %>%
  dplyr::select(age:or_time) %>%
  tbl_summary(
    by = sex,
    type = list(
      ven_res ~ "categorical",
      neo_chemo ~ "categorical",
      neo_xrt ~ "categorical"
    ),
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",    # drop min/max to save space
      all_categorical() ~ "{n} ({p}%)"),
    label = list(age ~ "Age (years)",
      race ~ "Race",
      liver_test ~ "Bilirubin Level (in mg/dL)", 
      op ~ "Type of Operation", 
      ebl ~ "Estimated Intraoperative Blood Loss (in mL)", 
      ven_res ~ "Resection + Reconstruction of Major Vasculature", 
      neo_chemo ~ "Chemotherapy Prior to Surgery", 
      neo_xrt ~ "Radiation Therapy Prior to Surgery", 
      les_size ~ "Tumor Size (in cm)", 
      or_time ~ "Operation Time (in mins)")) %>%
  add_p() %>%
  add_overall() %>%
  bold_labels() %>%
  modify_caption("**Baseline Characteristics by Sex**")
table1v2 <- table1 %>%
  as_gt() %>%
  gt::tab_options(
    data_row.padding = px(2),
    table.width      = pct(100),
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    heading.border.bottom.color = "black",
    table.font.color = "black",
    column_labels.border.bottom.style = "none", # Remove header line
    table_body.border.bottom.style = "none",    # Remove body line
    table_body.border.top.style = "none",
    table.border.top.style = "none",
    table.font.names = "arial",
    table.font.size = "12px" 
  ) |> 
  cols_width(
    label      ~ px(150),
    everything() ~ px(80)
  ) |> 
  gt::tab_style(
    style = gt::cell_borders(), 
    locations = gt::cells_body(columns = everything()))  |> 
  gt::gtsave(filename = "figures_tables/table1.pdf")
system("pdfcrop figures_tables/table1.pdf figures_tables/table1.pdf")

table1 |> 
  as_kable()
```

b) Grouped by Race
```{r}
table1byrace  <- pancreatic_df %>%
  dplyr::select(age:or_time) %>%
  tbl_summary(
    by = race,
    type = list(
      ven_res ~ "categorical",
      neo_chemo ~ "categorical",
      neo_xrt ~ "categorical"
    ),
    statistic = list(all_continuous()  ~ "{mean} ({sd})",    # drop min/max to save space
      all_categorical() ~ "{n} ({p}%)"),
    label = list(age ~ "Age (years)",
      sex ~ "Sex",
      liver_test ~ "Bilirubin Level (in mg/dL)", 
      op ~ "Type of Operation", 
      ebl ~ "Estimated Intraoperative Blood Loss (in mL)", 
      ven_res ~ "Resection + Reconstruction of Major Vasculature", 
      neo_chemo ~ "Chemotherapy Prior to Surgery", 
      neo_xrt ~ "Radiation Therapy Prior to Surgery", 
      les_size ~ "Tumor Size (in cm)", 
      or_time ~ "Operation Time (in mins)")) %>%
  add_p() %>%
  add_overall() %>%
  bold_labels() %>%
  modify_caption("**Baseline Characteristics by Race**")
table1bv2 <- table1byrace %>%
  as_gt() %>%
  gt::tab_options(
    data_row.padding = px(2),
    table.width      = pct(100),
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    heading.border.bottom.color = "black",
    table.font.color = "black",
    column_labels.border.bottom.style = "none", # Remove header line
    table_body.border.bottom.style = "none",    # Remove body line
    table_body.border.top.style = "none",
    table.border.top.style = "none",
    table.font.names = "arial",
    table.font.size = "12px" 
  ) |> 
  cols_width(
    label      ~ px(150),
    everything() ~ px(80)
  ) |> 
  gt::tab_style(
    style = gt::cell_borders(), 
    locations = gt::cells_body(columns = everything()))  |> 
  gt::gtsave(filename = "figures_tables/table1b.pdf")
system("pdfcrop figures_tables/table1b.pdf figures_tables/table1b.pdf")

table1byrace |> 
  as_kable()
```

###correlation between predictors and within covariates
```{r}
cor_dat <- pancreatic_df |> 
  dplyr::select(age, liver_test, ebl, les_size, or_time)
cor_matrix <- cor(cor_dat, method = "pearson")
rownames(cor_matrix) <- c("Age", "Bilirubin Level", "Estimated Blood Loss", "Tumor Size", "Operation Time")
colnames(cor_matrix) <- c("Age", "Bilirubin Level", "Estimated Blood Loss", "Tumor Size", "Operation Time")
pdf("figures_tables/corr_contvars.pdf", width = 6, height = 5)
corrplot(cor_matrix, method = "circle")
dev.off()
corrplot(cor_matrix, method = "circle")
```

Male
```{r}
cor_dat <- pancreatic_df |> 
  dplyr::filter(sex == "Male") |> 
  dplyr::select(age, liver_test, ebl, les_size, or_time)
cor_matrix <- cor(cor_dat, method = "pearson")
rownames(cor_matrix) <- c("Age", "Bilirubin Level", "Estimated Blood Loss", "Tumor Size", "Operation Time")
colnames(cor_matrix) <- c("Age", "Bilirubin Level", "Estimated Blood Loss", "Tumor Size", "Operation Time")
pdf("figures_tables/corr_contvars_male.pdf", width = 6, height = 5)
corrplot(cor_matrix, method = "circle")
dev.off()
corrplot(cor_matrix, method = "circle")
```

Female
```{r}
cor_dat <- pancreatic_df |> 
  dplyr::filter(sex == "Female") |> 
  dplyr::select(age, liver_test, ebl, les_size, or_time)
cor_matrix <- cor(cor_dat, method = "pearson")
rownames(cor_matrix) <- c("Age", "Bilirubin Level", "Estimated Blood Loss", "Tumor Size", "Operation Time")
colnames(cor_matrix) <- c("Age", "Bilirubin Level", "Estimated Blood Loss", "Tumor Size", "Operation Time")
pdf("figures_tables/corr_contvars_female.pdf", width = 6, height = 5)
corrplot(cor_matrix, method = "circle")
dev.off()
corrplot(cor_matrix, method = "circle")
```
operation time and age negatively correlated in men but no to low positive correlation in women

```{r}
df_filter <- pancreatic_df %>%
    group_by(sex) |> 
    mutate(age_quintile = ntile(age, 5)*20) |> 
  group_by(sex, age_quintile) |> 
  mutate(age_group = paste(min(age), max(age), sep = "-"))
  
plot <- ggplot(df_filter,
                 aes(x = or_time, y = ebl, color = sex, group = sex)) +
    geom_point(size = 0.5, alpha = 0.5) +
  geom_smooth() +
    labs (x = "Operation Time", y = "EBL", 
          color = "Sex", title = "Operation Time versus EBL by Age Quintile") + 
    facet_wrap(~age_quintile) +
  theme_classic()
pdf("figures_tables/ebl_vs_ortime_age.pdf", 
       width = 10, height = 8)
plot
dev.off()
```


### EBL Distribution in each categorical covariates
```{r, fig.height = 6, fig.width=8}
colnames <- c("op","ven_res", "neo_chemo", "neo_xrt", "race")
names <- c("Operation Type", "Reconstruction of Major Vasculature", "Prior Chemotherapy", "Prior Radiation Therapy", "Race")
plots <- lapply(1:length(colnames), function(x){
  plot <- ggplot(data = pancreatic_df |> drop_na(race),
                 aes(x = get(colnames[x]), y = ebl, fill = sex)) +
    geom_boxplot(alpha = 0.5) +
    labs (x = names[x], y = "Estimated Blood Loss", 
          fill = "Sex") +
    theme_classic()
  return(plot)
})

do.call(gridExtra::grid.arrange, c(plots, ncol = 3))

pdf("figures_tables/ebl_dist_across_cat_boxplot.pdf", 
       width = 8, height = 7)
do.call(gridExtra::grid.arrange, c(plots, ncol = 2))
dev.off()
```

```{r, fig.height = 6, fig.width=8}
colnames <- c("sex", "op","ven_res", "neo_chemo", "neo_xrt", "race")
names <- c("Sex", "Operation Type", "Reconstruction of Major Vasculature", "Prior Chemotherapy", "Prior Radiation Therapy", "Race")
plots <- lapply(1:length(colnames), function(x){
  plot <- ggplot(data = pancreatic_df |> drop_na(race),
                 aes(x = ebl, fill = get(colnames[x]))) +
    geom_density(alpha = 0.4) +
    labs (fill = names[x], x = "Estimated Blood Loss") +
    theme_classic() +
    theme(legend.position = "bottom")
  return(plot)
})

do.call(gridExtra::grid.arrange, c(plots, ncol = 3))

pdf("figures_tables/ebl_dist_across_cat_densityplot.pdf", 
       width = 8, height = 7)
do.call(gridExtra::grid.arrange, c(plots, ncol = 2))
dev.off()
```
EBL lower in female regardless of prior chemotherapy, radiation therapy and major reconstruction status. However, EBL in female vs male are different depending on operation type (effect modification probably, can try interaction term). Low EBL in white and hispanic women. Very few asian and black men recruited in the study so comparison is difficult. 

EBL's distribution is relatively normal across each group.

Recode race:
```{r}
pancreatic_df <- pancreatic_df |> 
  mutate(race = case_when(race == "White" ~ "White", 
                          race %in% c("Asian", "Black", "Hispanic") ~ "Non-white"), 
         race = fct_relevel(race, "White")) 
```

Modeling:
```{r}
lm_v0 <- lm(ebl ~ race*sex + age + op + ven_res + neo_chemo + 
              neo_xrt + liver_test + les_size + age*or_time,
            data = pancreatic_df)

summary(lm_v0)
```

Type of operation not really significant
```{r}
lm_v1 <- lm(ebl ~ race*sex + age + ven_res + neo_chemo + 
              neo_xrt + liver_test + les_size + age*or_time,
            data = pancreatic_df)

anova(lm_v1, lm_v0)
```
the model with operation time doesn't provide better fit. So removing that

```{r}
lm_v3 <- lm(ebl ~ race*sex + age + ven_res + neo_chemo + 
              neo_xrt + liver_test + les_size + age + or_time,
            data = pancreatic_df)
anova(lm_v3,lm_v1)
```
One with interactions for raceXsex, liver_testXles_size, ageXor_time significantly better

```{r}
pdf("figures_tables/lm_dharmaplot.pdf", width = 7, height = 5)
plot(lm_v1)
dev.off()
sim_ad <- scale(residuals(lm_v1))
ks.test(sim_ad, "pnorm")
```

```{r}
pdf("figures_tables/lm_diagnostic_plot.pdf",
       width = 5, height = 12)
par(mfrow=c(4,1), mar = c(2, 4, 2, 1))
plot(lm_v1)
par(mfrow=c(1, 1))
dev.off()
```

###Look at missingness
```{r}
table1byrace  <- pancreatic_df %>%
  dplyr::select(c(age:or_time, race2)) %>%
  tbl_summary(
    by = race2,
    type = list(
      ven_res ~ "categorical",
      neo_chemo ~ "categorical",
      neo_xrt ~ "categorical"
    ),
    statistic = list(all_continuous()  ~ "{mean} ({sd})",    # drop min/max to save space
      all_categorical() ~ "{n} ({p}%)"),
    label = list(age ~ "Age (years)",
      sex ~ "Sex",
      liver_test ~ "Bilirubin Level (in mg/dL)", 
      op ~ "Type of Operation", 
      ebl ~ "Estimated Intraoperative Blood Loss (in mL)", 
      ven_res ~ "Resection + Reconstruction of Major Vasculature", 
      neo_chemo ~ "Chemotherapy Prior to Surgery", 
      neo_xrt ~ "Radiation Therapy Prior to Surgery", 
      les_size ~ "Tumor Size (in cm)", 
      or_time ~ "Operation Time (in mins)")) %>%
  add_p() %>%
  add_overall() %>%
  bold_labels() %>%
  modify_caption("**Baseline Characteristics by Race**")
table1bv2 <- table1byrace %>%
  as_gt() %>%
  gt::tab_options(
    data_row.padding = px(2),
    table.width      = pct(100),
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    heading.border.bottom.color = "black",
    table.font.color = "black",
    column_labels.border.bottom.style = "none", # Remove header line
    table_body.border.bottom.style = "none",    # Remove body line
    table_body.border.top.style = "none",
    table.border.top.style = "none",
    table.font.names = "arial",
    table.font.size = "12px" 
  ) |> 
  cols_width(
    label      ~ px(150),
    everything() ~ px(80)
  ) |> 
  gt::tab_style(
    style = gt::cell_borders(), 
    locations = gt::cells_body(columns = everything()))  |> 
  gt::gtsave(filename = "figures_tables/table_missingness.pdf")
system("pdfcrop figures_tables/table_missingness.pdf figures_tables/table_missingness.pdf")

table1byrace |> 
  as_kable()
```

```{r}
fig2b <- vis_miss(pancreatic_df |> dplyr::select(-prot_energy)) + 
  theme(axis.text.x = element_text(vjust = 0))
ggsave("figures_tables/missingness_info.png", fig2b, height = 7, width =10)
```

Missingness by gender and age
```{r, fig.height = 3, fig.width=6, warning=FALSE}
race_gender_missing <- pancreatic_df |>
  dplyr::select(-prot_energy) |> 
  group_by(sex) |> 
  mutate(total = length(unique(patid))) |> 
  ungroup() |> 
  drop_na(race) |> 
  group_by(sex) |> 
  summarize (missing_pct = (total[1] - length(unique(patid)))*100/total[1]) |> 
  mutate(dummy_group = 1)

missing_by_gender <- ggplot(aes(x = sex, y = missing_pct, group = dummy_group, color = sex), 
       data = race_gender_missing) +
  geom_point() +
  geom_line() +
  geom_line() +
  labs(x = "Sex", y = "Missing Data %") +
  theme_classic()

ggsave("figures_tables/missingness_byracegender.pdf", missing_by_gender, height = 3, width =5)
system("pdfcrop figures_tables/missingness_byracegender.pdf figures_tables/missingness_byracegender.pdf")
missing_by_gender
```

Look for other possible sources. 
- check trend of ebl across each continuous variable to see if its even linear or not, if not linear can try gams

Missingness in different quantile of age or other continuous variables
```{r}
colnames <- c("sex", "op","ven_res", "neo_chemo", "neo_xrt")
names <- c("Sex", "Operation Type", "Reconstruction of Major Vasculature", "Prior Chemotherapy", "Prior Radiation Therapy")
categorical_missing_pct <- lapply(1:length(colnames), function(x){
  race_gender_missing <- pancreatic_df |>
    dplyr::select(-prot_energy) |> 
    group_by(get(colnames[x])) |> 
    mutate(total = length(unique(patid))) |> 
    ungroup() |> 
    drop_na(race) |> 
    group_by(get(colnames[x])) |> 
    summarize (missing_pct = (total[1] - length(unique(patid)))*100/total[1]) |> 
    mutate(dummy_group = 1)
  colnames(race_gender_missing) <- c(colnames[x], "missing_pct", "dummy_group")
  print(head(race_gender_missing))
  
  missing_by_cat <- ggplot(aes(x = get(colnames[x]), y = missing_pct, group = dummy_group), 
         data = race_gender_missing) +
    geom_point() +
    geom_line() +
    geom_line() +
    labs(x = names[x], y = "Missing Data %") +
    theme_classic()
  
  return(missing_by_cat)
})

do.call(gridExtra::grid.arrange, c(categorical_missing_pct, ncol = 3))

pdf("figures_tables/missingness_percategorical_group.pdf", 
       width = 9, height = 8)
do.call(gridExtra::grid.arrange, c(categorical_missing_pct, ncol = 3))
dev.off()
```

```{r}
colnames <- c("sex", "op","ven_res", "neo_chemo", "neo_xrt")
names <- c("Sex", "Operation Type", "Reconstruction of Major Vasculature", "Prior Chemotherapy", "Prior Radiation Therapy")
categorical_missing_pct <- lapply(1:length(colnames), function(x){
  a <- pancreatic_df |> 
    mutate(xcol = get(colnames[x])) |> 
    group_by(xcol, race) |> 
    summarise(n = n()) |> 
    mutate(sum = sum(n)) |> 
    dplyr::filter(is.na(race) == TRUE) |> 
    mutate(yes = n/sum, no = 1- yes) |> 
    dplyr::select(xcol, n, sum, yes, no) |> 
    ungroup() |> pivot_longer(cols = yes:no)
  
  missing_by_cat <- ggplot(aes(x = xcol, y = value, fill = name), 
         data = a) +
    geom_bar(stat = "identity", position = "fill") +
    labs(x = names[x], fill = "Missing Data", y = "") +
    theme_classic()
  
  return(missing_by_cat)
})

do.call(gridExtra::grid.arrange, c(categorical_missing_pct, ncol = 3))

pdf("figures_tables/missingness_percategorical_group_barplot.pdf", 
       width = 11, height = 7)
do.call(gridExtra::grid.arrange, c(categorical_missing_pct, ncol = 3))
dev.off()
```

```{r}
colnames <- c("ebl", "age","liver_test", "les_size", "or_time")
names <- c("Estimated Blood Loss", "Age", "Bilirubin Level (in mg/dL)", "Tumor Size (in cm)", "Operation Time (in minutes)")
get_quintile_missing_pct <- lapply(1:length(colnames), function(x){
  df_filter <- pancreatic_df %>%
    group_by(sex) |> 
    mutate(xcol = get(colnames[x]), 
           x_percentile = ntile(xcol, 20)*5) |> 
    group_by(x_percentile, sex) |> 
    summarise(n = n(), 
              missing_y = sum(is.na(race)),
              group = paste(as.character(min(xcol)), as.character(max(xcol)), sep = "-"),
              covariate = colnames[x],
              pct_missing_y = mean(is.na(race)) * 100)
  
  plot <- ggplot(df_filter,
                 aes(x = x_percentile, y = pct_missing_y)) +
    geom_point(size = 0.5, alpha = 0.5) +
    labs (x = paste(names[x], "(Percentile)" ), y = "% of Missing Race Info", 
          color = "Sex") + 
  theme_classic() +
    theme(legend.position = "none")
  
  return(plot)
})

do.call(gridExtra::grid.arrange, c(get_quintile_missing_pct, ncol = 2))

pdf("figures_tables/missingness_percontinuous_group.pdf", 
       width = 10, height = 8)
do.call(gridExtra::grid.arrange, c(get_quintile_missing_pct, ncol = 2))
dev.off()
```
The distribution is missing percent against all the continous variables are randomly distributed.

```{r}
mcar_test(pancreatic_df |>  dplyr::select(-prot_energy))
```
significant, reject MCAR, moving forward with MAR assumption

Tumor size (heavily left-skewed)
```{r}
colnames <- c("ebl","age","liver_test", "les_size", "or_time")
names <- c("Estimated Blood Loss", "Age", "Bilirubin Level (in mg/dL)", "Tumor Size (in cm)", "Operation Time (in minutes)")
plots <- lapply(1:length(colnames), function(x){
  plot <- ggplot(data = pancreatic_df |> drop_na(race),
                 aes(x = get(colnames[x]))) +
    geom_density(alpha = 0.5) +
    labs (x = names[x]) +
    theme_classic()
  return(plot)
})

do.call(gridExtra::grid.arrange, c(plots, ncol = 3))

pdf("figures_tables/cont_variables_distribution.pdf", 
       width = 8, height = 6)
do.call(gridExtra::grid.arrange, c(plots, ncol = 3))
dev.off()
```

```{r}
outliers_filtered <- pancreatic_df |> 
  drop_na(race) |> 
  dplyr::filter(les_size < 11, 
                liver_test < 20)
  
  
colnames <- c("age","liver_test", "les_size", "or_time")
names <- c("Age", "Bilirubin Level (in mg/dL)", "Tumor Size (in cm)", "Operation Time (in minutes)")
plots <- lapply(1:length(colnames), function(x){
  plot <- ggplot(data = outliers_filtered,
                 aes(x = get(colnames[x]), y = ebl)) +
    geom_point(size = 0.8) +
    geom_smooth() +
    labs (x = names[x], y = "Estimated Blood Loss") +
    theme_classic()
  return(plot)
})

do.call(gridExtra::grid.arrange, c(plots, ncol = 2))

pdf("figures_tables/ebl_versus_cont_variable.pdf", 
       width = 8, height = 6)
do.call(gridExtra::grid.arrange, c(plots, ncol = 2))
dev.off()
```
Bilirubin level - minimal association at smaller amount, increased association after 5

GAM

```{r}
gam_v1 <- gam(ebl ~ race*sex + age +  sex + ven_res + neo_chemo + 
              neo_xrt + s(liver_test) + les_size + age*or_time,
            data = pancreatic_df)

summary(gam_v1)
gam.check(gam_v1)
```

```{r}
gam_v2 <- gam(ebl ~ race*sex + age +  sex + ven_res + neo_chemo + 
              neo_xrt + s(liver_test) + les_size + or_time,
            data = pancreatic_df)

summary(gam_v2)
anova(gam_v1, gam_v2)
```

selected gam_v2
```{r}
pdf("figures_tables/gamv1_diagnostic_plot.pdf",
       width = 5, height = 12)
par(mfrow=c(4,1), mar = c(2, 4, 2, 1))
gam.check(gam_v2)
par(mfrow=c(1, 1))
dev.off()
```

```{r}
sim_ad <- scale(residuals(gam_v2))
ks.test(sim_ad, "pnorm")
```


###Compare fit
```{r}
allmods_diag <- data.frame(
  Model = c("LM", "GAM"),
  AIC   = c(AIC(lm_v1), AIC(gam_v2)),
  BIC   = c(BIC(lm_v1), BIC(gam_v2)),
  `log(Likelihood)` = c(as.numeric(logLik(lm_v1)), as.numeric(logLik(gam_v2))), 
  DF     = c(attr(logLik(lm_v1), "df"),
             attr(logLik(gam_v2), "df"))
) |> 
  kable(caption = "", 
        booktabs = FALSE, longtable = FALSE, linesep = "", digits = 2, full_width = FALSE)  |> 
  kable_styling(latex_options = "hold_position", font_size = 12) 

#save_kable(allmods_diag, file = "figures_tables/model_comparison.pdf")
#system("pdfcrop figures_tables/model_comparison.pdf figures_tables/model_comparison.pdf")

allmods_diag
```

Based on this, lm_v1 is chosen 
```{r, fig.align='center', fig.width=5, fig.height=4, warning = FALSE, message = FALSE}
pdf("figures_tables/gamv1_smoothfunc.pdf", width = 8, height = 8)
plot(gam_v2, shade = TRUE, pages = 1)
dev.off()
#plot(lm_v1, shade = TRUE, pages = 1)
```

```{r, fig.align='center', fig.width=7, fig.height=4, warning = FALSE, message = FALSE}
plot_model(lm_v1, type = "pred", terms = c("or_time", "sex", "race")) +
  labs(title = "", 
       y = "Estimated Blood Loss (in mL) Prediction",
     x = "Operation Time (in minutes)") +
  theme_classic()
ggsave("figures_tables/ebl_pred_race_gender_operatn_time.pdf", height = 4, width =8)
```

```{r, fig.align='center', fig.width=8, fig.height=4, warning = FALSE, message = FALSE}
plot_model(lm_v1, type = "pred", terms = c("les_size", "ven_res", "sex")) + 
  labs(title = "", 
       y = "Estimated Blood Loss (in mL) Prediction",
     x = "Tumor Size (in cm)", color = "Major Vasculature Reconstruction") +
  theme_classic()
ggsave("figures_tables/ebl_pred_gender_ven_res_tumorsize.pdf", height = 4, width =10)
```
Same trend in men and women across all 4 races

```{r, fig.align='center', fig.width=4, fig.height=4, warning = FALSE, message = FALSE}
plot_model(lm_v1, type = "pred", terms = c("age", "sex", "race"), dodge = 0.6) +
  labs(title = "", 
       y = "Estimated Blood Loss (in mL) Prediction",
     x = "Age") +
  theme_classic() 
ggsave("figures_tables/ebl_pred_race_gender_age.pdf", height = 4, width =8)
```

```{r, fig.align='center', fig.width=5, fig.height=4, warning = FALSE, message = FALSE}
plot_mod <- plot_model(lm_v1, type = "pred", 
                       terms = c("or_time","neo_chemo", "neo_xrt"), dodge = 0.6) +
  labs(title = "", 
       y = "Estimated Blood Loss (in mL) Prediction",
     x = "Operation Time", color = "Prior Chemotherapy") +
  theme_classic()

plot_mod$data$facet <- c("Radiation Therapy: No", "Radiation Therapy: Yes")
plot_mod
ggsave("figures_tables/ebl_pred_chem_rad_therapy.pdf", height = 4, width =8)
```

Consider that normal bilirubin level is 0.2-1.3. Compare the normal state with degree of increase or decrease from there
```{r, fig.align='center', fig.width=4, fig.height=4, warning = FALSE, message = FALSE}
plot_model(lm_v1, type = "pred", terms = c("sex", "race"), dodge = 0.6) +
  labs(title = "", 
       y = "Estimated Blood Loss (in mL) Prediction",
     x = "Age") +
  theme_classic() 
ggsave("figures_tables/ebl_pred_race_gender.pdf", height = 4, width =8)
```

\newpage
##results table

LM
```{r}
bc_tbl <- tbl_regression(
  lm_v1,
  intercept= TRUE,
  conf.int = TRUE,
  p.value = TRUE,
  include = everything(),
  tidy_fun = broom.helpers::tidy_parameters,
  estimate_fun = function(x) sprintf("%.5f", x),
    label = list(age ~ "Age (years)",
                 race ~ "Race",
      sex ~ "Sex",
      `liver_test` ~ "Bilirubin Level (in mg/dL)", 
      ven_res ~ "Resection + Reconstruction of Major Vasculature", 
      neo_chemo ~ "Chemotherapy Prior to Surgery", 
      neo_xrt ~ "Radiation Therapy Prior to Surgery", 
      `les_size` ~ "Tumor Size (in cm)", 
      or_time ~ "Operation Time (in mins)"),
) %>%
  bold_labels() %>%  
  modify_header(label ~ "Term") 

bc_tbl |> 
  as_gt() |> 
  gt::tab_options(
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    heading.border.bottom.color = "black",
    table.font.color = "black",
    column_labels.border.bottom.style = "none", # Remove header line
    table_body.border.bottom.style = "none",    # Remove body line
    table_body.border.top.style = "none",
    table.border.top.style = "none",
    table.font.names = "arial",
    table.font.size = "12px" 
  ) |> 
  gt::tab_style(
    style = gt::cell_borders(), 
    locations = gt::cells_body(columns = everything())) |> gt::gtsave("figures_tables/lm_res_tab.pdf")

system("pdfcrop figures_tables/lm_res_tab.pdf figures_tables/lm_res_tab.pdf")

bc_tbl |> 
  as_kable()
```

GAM
```{r}
bc_tbl <- tbl_regression(
  gam_v2,
  intercept= TRUE,
  conf.int = TRUE,
  p.value = TRUE,
  include = everything(),
  tidy_fun = broom.helpers::tidy_parameters,
  estimate_fun = function(x) sprintf("%.5f", x),
    label = list(age ~ "Age (years)",
                 race ~ "Race",
      sex ~ "Sex",
      `s(liver_test)` ~ "s(Bilirubin Level (in mg/dL))", 
      ven_res ~ "Resection + Reconstruction of Major Vasculature", 
      neo_chemo ~ "Chemotherapy Prior to Surgery", 
      neo_xrt ~ "Radiation Therapy Prior to Surgery", 
      `les_size` ~ "Tumor Size (in cm)", 
      or_time ~ "Operation Time (in mins)"),
) %>%
  bold_labels() %>%  
  modify_header(label ~ "Term") 

bc_tbl |> 
  as_gt() |> 
  gt::tab_options(
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    heading.border.bottom.color = "black",
    table.font.color = "black",
    column_labels.border.bottom.style = "none", # Remove header line
    table_body.border.bottom.style = "none",    # Remove body line
    table_body.border.top.style = "none",
    table.border.top.style = "none",
    table.font.names = "arial",
    table.font.size = "12px" 
  ) |> 
  gt::tab_style(
    style = gt::cell_borders(), 
    locations = gt::cells_body(columns = everything())) |> gt::gtsave("figures_tables/gam_res_tab.pdf")

system("pdfcrop figures_tables/gam_res_tab.pdf figures_tables/gam_res_tab.pdf")

bc_tbl |> 
  as_kable(digits = 3)
```

```{r}
## smooth terms part

smooth_df <- as.data.frame(summary(gam_v2)$s.table)
smooth_df$term <- rownames(smooth_df)

smooth_df <- smooth_df[, c("term", "edf", "F", "p-value")]
colnames(smooth_df) <- c("term", "edf", "statistic", "p.value")

smooth_df <-  smooth_df %>% mutate(p.value = case_when(p.value==0~"<0.001", TRUE~as.character(p.value))) |> 
  mutate(term = case_when(term == "s(liver_test)" ~ "s(Bilirubin Level (in mg/dL))"))


smooth_df %>%
  gt() |> 
  gt::tab_options(
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    heading.border.bottom.color = "black",
    table.font.color = "black",
    column_labels.border.bottom.style = "none", # Remove header line
    table_body.border.bottom.style = "none",    # Remove body line
    table_body.border.top.style = "none",
    table.border.top.style = "none",
    table.font.names = "arial",
    table.font.size = "12px" 
  ) |> 
  gt::tab_style(
    style = gt::cell_borders(), 
    locations = gt::cells_body(columns = everything())) |> 
  tab_header(title = "Smooth Terms") |> 
  gt::gtsave("figures_tables/gam_smooth_eff.pdf")

system("pdfcrop figures_tables/gam_smooth_eff.pdf figures_tables/gam_smooth_eff.pdf")
```

## sensitivity analysis (PMM)

lm_v1 for ebl with all other covariates, excluding race (for NAs and non-NAs)
```{r}
data_complete <- pancreatic_df |> 
  dplyr::filter(race2 != "NA")
data_incomplete <- pancreatic_df |> 
  dplyr::filter(race2== "NA")

lm_v1_complete <- lm(ebl ~ sex + age + ven_res + neo_chemo + 
              neo_xrt + liver_test + les_size + age*or_time,
            data = data_complete)
lm_v1_incomplete <- lm(ebl ~ sex + age + ven_res + neo_chemo + 
              neo_xrt + liver_test + les_size + age*or_time,
            data = data_incomplete)
lm_list <- list(lm_v1_complete, 
                lm_v1_incomplete)
names(lm_list) <- c("LM Complete", "LM Missing Race")
```

```{r}
pairs <- combn(names(lm_list), 2, simplify = FALSE)

lapply(pairs, function(compare){
  first_fit <- lm_list[[compare[1]]]
  second_fit <- lm_list[[compare[2]]]
  coef_first <- coef(first_fit)
  coef_second <- coef(second_fit)
  
  vcov_first <- vcov(first_fit)
  vcov_second <- vcov(second_fit)
  
  overlapping_terms <- intersect(names(coef_first), 
                                 names(coef_second))
  res_complete_late <- lapply(overlapping_terms, function(term){
  
    beta1 <- coef_first[term]
    beta2 <- coef_second[term]
  
    var1 <- vcov_first[term, term]
    var2 <- vcov_second[term, term]
  
    z <- (beta1 - beta2) / sqrt(var1 + var2)
    p <- 2 * (1 - pnorm(abs(z)))
    
    data.frame(
      `Group` = paste0(compare[1], " vs ", compare[2]), 
      `Beta1` = beta1, `Beta2` = beta2, `z` = z, `P-value` = p
    )
  })
  df = as.data.frame(do.call(rbind, res_complete_late))
  nrows = nrow(df)
  comp <- knitr::kable(df, digits = 3) |> 
    kable_styling(latex_options = c("striped", "hold_position"), font_size = 12)  |>
  row_spec(0, extra_css = "border: 1px solid black;") |>
  column_spec(1:6, 
              extra_css = "border-left: 1px solid black; border-right: 1px solid black;") |>
  row_spec(1:nrows, extra_css = "border-top: 1px solid black; border-bottom: 1px solid black;")
  print(comp)
  #save_kable(comp, file = paste0("figures_tables/coef_comparison", compare[1], compare[2], ".pdf"))
})
```
No significant difference in coefficients so results from CCA is robust. 

Sample-size calculation
```{r}
sig = 0.05
power = 0.80
d = 40/(summary(lm_v1)$sigma)
z_1minusa = qnorm(1-sig/2)
z_power = qnorm(power)
n_pergroup = (z_1minusa + z_power)^2/(d^2*0.25)
n_pergroup
nadj = n_pergroup/(1 - 0.2)
```


