---
title: "q1_analysis"
author: "Chhiring Lama"
date: "2026-04-27"
output:
  pdf_document:
    toc: yes
    toc_depth: 2
header-includes:
  - \usepackage{fancyhdr}
  - \usepackage{lipsum}
  - \pagestyle{fancy}
  - \fancyhead[R]{\thepage}
  - \fancypagestyle{plain}{\pagestyle{fancy}}
  - \usepackage{enumitem}
  - \usepackage{float}
  - \usepackage{amsmath}
  - \usepackage{amssymb}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gtsummary)
library(gt)
library(corrplot)
library(caret)
library(naniar)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

options(scipen = 999)
```

### Import and clean dataset
```{r}
pancreatic_df <- read_csv("data/ebl.csv") |> 
  janitor::clean_names() |> 
  mutate(sex = case_when(sex == 0 ~ "Male", 
                         sex == 1 ~ "Female"), 
         sex = fct_relevel(sex, "Male"),
         race = case_when(race == 0 ~ "White", 
                         race == 1 ~ "Black", 
                         race == 2 ~ "Asian", 
                         race == 3 ~ "Hispanic"), 
         race = fct_relevel(race, "White"), 
         op = as.factor(case_when(op == 0 ~ "Body or tail", 
                        op == 1 ~ "Whipple operation")),
         ven_res = as.factor(case_when(ven_res == 0 ~ "No", 
                             ven_res == 1 ~ "Yes")), 
         neo_chemo = as.factor(case_when(neo_chemo == 0 ~ "No", 
                             neo_chemo == 1 ~ "Yes")), 
         neo_xrt = as.factor(case_when(neo_xrt == 0 ~ "No", 
                             neo_xrt == 1 ~ "Yes")))

head(pancreatic_df)
```

table 1: demographic characteristics + others
```{r}
table1  <- pancreatic_df %>%
  dplyr::select(age:or_time) %>%
  tbl_summary(
    by = sex,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",    # drop min/max to save space
      all_categorical() ~ "{n} ({p}%)"
    ),
    label = list(
      age ~ "Age (years)",
      race ~ "Race",
      liver_test ~ "Bilirubin Level (in mg/dL)", 
      op ~ "Type of Operation", 
      ebl ~ "Estimated Intraoperative Blood Loss (in mL)", 
      ven_res ~ "Resection + Reconstruction of Major Vasculature", 
      neo_chemo ~ "Chemotherapy Prior to Surgery", 
      neo_xrt ~ "Radiation Therapy Prior to Surgery", 
      les_size ~ "Tumor Size (in cm)", 
      or_time ~ "Operation Time (in mins)"
    )
  ) %>%
  add_p() %>%
  add_overall() %>%
  bold_labels() %>%
  modify_caption("**Baseline Characteristics by Sex**")
table1v2 <- table1 %>%
  as_gt() %>%
  gt::tab_options(
    data_row.padding = px(2),
    table.width      = pct(100),
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    heading.border.bottom.color = "black",
    table.font.color = "black",
    column_labels.border.bottom.style = "none", # Remove header line
    table_body.border.bottom.style = "none",    # Remove body line
    table_body.border.top.style = "none",
    table.border.top.style = "none",
    table.font.names = "arial",
    table.font.size = "12px" 
  ) |> 
  cols_width(
    label      ~ px(150),
    everything() ~ px(80)
  ) |> 
  gt::tab_style(
    style = gt::cell_borders(), 
    locations = gt::cells_body(columns = everything()))  |> 
  gt::gtsave(filename = "figures_tables/table1.pdf")
system("pdfcrop figures_tables/table1.pdf figures_tables/table1.pdf")

table1 |> 
  as_kable()
```

```{r}
table1byrace  <- pancreatic_df %>%
  dplyr::select(age:or_time) %>%
  tbl_summary(
    by = race,
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",    # drop min/max to save space
      all_categorical() ~ "{n} ({p}%)"
    ),
    label = list(
      age ~ "Age (years)",
      sex ~ "Sex",
      liver_test ~ "Bilirubin Level (in mg/dL)", 
      op ~ "Type of Operation", 
      ebl ~ "Estimated Intraoperative Blood Loss (in mL)", 
      ven_res ~ "Resection + Reconstruction of Major Vasculature", 
      neo_chemo ~ "Chemotherapy Prior to Surgery", 
      neo_xrt ~ "Radiation Therapy Prior to Surgery", 
      les_size ~ "Tumor Size (in cm)", 
      or_time ~ "Operation Time (in mins)"
    )
  ) %>%
  add_p() %>%
  add_overall() %>%
  bold_labels() %>%
  modify_caption("**Baseline Characteristics by Sex**")
table1bv2 <- table1byrace %>%
  as_gt() %>%
  gt::tab_options(
    data_row.padding = px(2),
    table.width      = pct(100),
    table.border.top.color = "black",
    table.border.bottom.color = "black",
    heading.border.bottom.color = "black",
    table.font.color = "black",
    column_labels.border.bottom.style = "none", # Remove header line
    table_body.border.bottom.style = "none",    # Remove body line
    table_body.border.top.style = "none",
    table.border.top.style = "none",
    table.font.names = "arial",
    table.font.size = "12px" 
  ) |> 
  cols_width(
    label      ~ px(150),
    everything() ~ px(80)
  ) |> 
  gt::tab_style(
    style = gt::cell_borders(), 
    locations = gt::cells_body(columns = everything()))  |> 
  gt::gtsave(filename = "figures_tables/table1b.pdf")
system("pdfcrop figures_tables/table1b.pdf figures_tables/table1b.pdf")

table1byrace |> 
  as_kable()
```

###correlation between predictors and within covariates

Male
```{r}
cor_dat <- pancreatic_df |> 
  dplyr::filter(sex == "Male") |> 
  dplyr::select(age, liver_test, ebl, les_size, or_time)
cor_matrix <- cor(cor_dat, method = "pearson")
rownames(cor_matrix) <- c("Age", "Bilirubin Level", "Estimated Blood Loss", "Tumor Size", "Operation Time")
colnames(cor_matrix) <- c("Age", "Bilirubin Level", "Estimated Blood Loss", "Tumor Size", "Operation Time")
pdf("figures_tables/corr_contvars_male.pdf", width = 6, height = 5)
corrplot(cor_matrix, method = "circle")
dev.off()
corrplot(cor_matrix, method = "circle")
```

Female
```{r}
cor_dat <- pancreatic_df |> 
  dplyr::filter(sex == "Female") |> 
  dplyr::select(age, liver_test, ebl, les_size, or_time)
cor_matrix <- cor(cor_dat, method = "pearson")
rownames(cor_matrix) <- c("Age", "Bilirubin Level", "Estimated Blood Loss", "Tumor Size", "Operation Time")
colnames(cor_matrix) <- c("Age", "Bilirubin Level", "Estimated Blood Loss", "Tumor Size", "Operation Time")
pdf("figures_tables/corr_contvars_female.pdf", width = 6, height = 5)
corrplot(cor_matrix, method = "circle")
dev.off()
corrplot(cor_matrix, method = "circle")
```
operation time and age negatively correlated in men but no to low positive correlation in women

### EBL Distribution in each categorical covariates
```{r, fig.height = 6, fig.width=10}
colnames <- c("op","ven_res", "neo_chemo", "neo_xrt", "race")
names <- c("Operation Type", "Reconstruction of Major Vasculature", "Prior Chemotherapy", "Prior Radiation Therapy", "Race")
plots <- lapply(1:length(colnames), function(x){
  plot <- ggplot(data = pancreatic_df |> drop_na(race),
                 aes(x = get(colnames[x]), y = ebl, fill = sex)) +
    geom_boxplot(alpha = 0.5) +
    labs (x = names[x], y = "Estimated Blood Loss", 
          fill = "Sex") +
    theme_classic()
  return(plot)
})

do.call(gridExtra::grid.arrange, c(plots, ncol = 3))

pdf("figures_tables/ebl_dist_across_cat_boxplot.pdf", 
       width = 8, height = 7)
do.call(gridExtra::grid.arrange, c(plots, ncol = 2))
dev.off()
```

```{r, fig.height = 6, fig.width=10}
colnames <- c("sex", "op","ven_res", "neo_chemo", "neo_xrt", "race")
names <- c("Sex", "Operation Type", "Reconstruction of Major Vasculature", "Prior Chemotherapy", "Prior Radiation Therapy", "Race")
plots <- lapply(1:length(colnames), function(x){
  plot <- ggplot(data = pancreatic_df |> drop_na(race),
                 aes(x = ebl, fill = get(colnames[x]))) +
    geom_density(alpha = 0.4) +
    labs (fill = names[x], x = "Estimated Blood Loss") +
    theme_classic() +
    theme(legend.position = "bottom")
  return(plot)
})

do.call(gridExtra::grid.arrange, c(plots, ncol = 3))

pdf("figures_tables/ebl_dist_across_cat_densityplot.pdf", 
       width = 8, height = 7)
do.call(gridExtra::grid.arrange, c(plots, ncol = 2))
dev.off()
```
EBL lower in female regardless of prior chemotherapy, radiation therapy and major reconstruction status. However, EBL in female vs male are different depending on operation type (effect modification probably, can try interaction term). Low EBL in white and hispanic women. Very few asian and black men recruited in the study so comparison is difficult. 

EBL's distribution is relatively normal across each group.

Modeling:
```{r}
lm_v1 <- lm(ebl ~ race*sex + age +  sex*op + ven_res + neo_chemo + 
              neo_xrt + liver_test*les_size + age*or_time,
            data = pancreatic_df)

summary(lm_v1)
```

###Look at missingness
```{r}
fig2b <- vis_miss(pancreatic_df |> dplyr::select(-prot_energy)) + 
  theme(axis.text.x = element_text(vjust = 0))
ggsave("figures_tables/missingness_info.png", fig2b, height = 7, width =10)
```

Missingness by gender and age
```{r, fig.height = 3, fig.width=6, warning=FALSE}
race_gender_missing <- pancreatic_df |>
  dplyr::select(-prot_energy) |> 
  group_by(sex) |> 
  mutate(total = length(unique(patid))) |> 
  ungroup() |> 
  drop_na(race) |> 
  group_by(sex) |> 
  summarize (missing_pct = (total[1] - length(unique(patid)))*100/total[1]) |> 
  mutate(dummy_group = 1)

missing_by_gender <- ggplot(aes(x = sex, y = missing_pct, group = dummy_group, color = sex), 
       data = race_gender_missing) +
  geom_point() +
  geom_line() +
  geom_line() +
  labs(x = "Sex", y = "Missing Data %") +
  theme_classic()

ggsave("figures_tables/missingness_byracegender.pdf", missing_by_gender, height = 3, width =5)
system("pdfcrop figures_tables/missingness_byracegender.pdf figures_tables/missingness_byracegender.pdf")
missing_by_gender
```

Look for other possible sources
